Semi-supervised author disambiguation of large-scale collections of labeled scientific publications

### 2015-08-05 paper.tex
save

\title{Semi-supervised author disambiguation of scientific publications}

@@

In academic digital libraries, author name disambiguation is the problem of
grouping together publications written by a same person.  Author name
disambiguation is often a difficult problem because an author may use different
spellings, variants or synonyms across her career and/or distinct authors may
share the same name. Most notably, author disambiguation is often even more
troublesome in the case of researchers from non-Western cultures, where
personal names may be traditionally less diverse (leading to homonym issues) or
for which transliteration to latin characters may not be unique (leading to
synonym issues). With the fast growth of the scientific literature, author
disambiguation has become a pressing issue because the accuracy of information
managed at the level of individuals directly affects the relevance search
results (e.g., when querying for all publications written by a given author),
the reliability of bibliometrics (e.g., citation counts or other impact
metrics) or the relevance of scientific network analysis.

Efforts and solutions to author disambiguation have been proposed from various
communities \citep{liu2014author}. On the one hand, libraries have maintained
authorship control through manual curation, either in a centralized way by
hiring collaborators or by developing services inviting authors to register
their publications themselves (e.g., Google Scholar, Inspire-HEP, ...). Recently,
efforts also started to create persistent digital identifiers assigned to researchers
(e.g., ORCID, ResearchedID, ...), with the further objective to embed these
identifiers in the submission workflow of publishers or repositeries (e.g.,
Elsevier, arXiv, Inspire-HEP, ...), thereby univocally solving any disambiguation
issue. With the large cost of centralized manual authorship control or until
the larger adoption of crowdsourced solutions, the impact of these efforts are
unfortunately limited by the efficiency, motivation or integrity of their
active contributors. Similarly, the success of persistent digital identifier
efforts is conditioned to a large and ubiquituous adoption by both researchers and publishers.
For these reasons, fully automated machine learning-based methods have been
proposed during the past decade \glnote{Add references} to provide immediate, less costly
and efficient solutions to author disambiguation. In this work, our goal
is to explore and demonstrate how both approaches can coexist and benefit from
each other.  In particular, we study how labeled data obtained through manual
curation (either centralized or crowdsourced) can be exploited (i) to learn an
accurate classifier for identifying corefering authors, (ii) to guide the
clustering of scientific publications by distinct authors in a semi-supervised
way and finally (iii) to match the resulting clusters with existing profiles.
Our analysis of parameters and features on this large training set reveal
that... \glnote{To be completed with actual results.}

@@

\glnote{I will take care of this}

@@

% Before presenting the methodology, we need to define some terminologies we will
% use to describe the data entities. An author's full name, or author name, is
% the string variation that was mention in the database record to refer to the
% author. The full name is a composition of the author last name and the author
% other names. For example, in the full name (Kerth, Leroy T.), the last name is
% Kerth and the other names will be (Leroy T.). We also use the terminology
% initials. In our full name example, the initials will be (L T). First initial
% then will be (L). So, grouping publications  What we refer to as the real
% author is the real person behind the publication. We will use the term
% signature as the entity that we apply clustering on.  A signature is a unique
% combination between a publication and an author name as mentioned in the
% database. The term blocking means grouping the signature by a certain version
% of the author name. For example, blocking by the author last name and first
% name initial means grouping together all signatures that have in common the
% same author's last name and first name initial. In our example, grouping by the
% string (Kerth L). What we mean by claimed publications is the publications in
% which their real author identity were claimed. We use these claimed
% publications to generate our ground truth data.

>>>>

\title{Semi-supervised author disambiguation of large-scale collections of labeled scientific publications}

@@

In academic digital libraries, author name disambiguation is the problem of
grouping together publications written by a same person.  Author name
disambiguation is often a difficult problem because an author may use different
spellings, variants or synonyms across her career and/or distinct authors may
share the same name. Most notably, author disambiguation is often even more
troublesome in the case of researchers from non-Western cultures, where
personal names may be traditionally less diverse (leading to homonym issues) or
for which transliteration to latin characters may not be unique (leading to
synonym issues). With the fast growth of the scientific literature, author
disambiguation has become a pressing issue because the accuracy of information
managed at the level of individuals directly affects the relevance search
results (e.g., when querying for all publications written by a given author),
the reliability of bibliometrics and author rankings (e.g., citation counts or other impact
metrics, as studied in \citep{strotmann2012author}) or the relevance of scientific network analysis.

Efforts and solutions to author disambiguation have been proposed from various
communities \citep{liu2014author}. On the one hand, libraries have maintained
authorship control through manual curation, either in a centralized way by
hiring collaborators or by developing services inviting authors to register
their publications themselves (e.g., Google Scholar, Inspire-HEP, ...). Recently,
efforts also started to create persistent digital identifiers assigned to researchers
(e.g., ORCID, ResearchedID, ...), with the further objective to embed these
identifiers in the submission workflow of publishers or repositeries (e.g.,
Elsevier, arXiv, Inspire-HEP, ...), thereby univocally solving any disambiguation
issue. With the large cost of centralized manual authorship control or until
the larger adoption of crowdsourced solutions, the impact of these efforts are
unfortunately limited by the efficiency, motivation or integrity of their
active contributors. Similarly, the success of persistent digital identifier
efforts is conditioned to a large and ubiquituous adoption by both researchers and publishers.
For these reasons, fully automated machine learning-based methods have been
proposed during the past decade (See Section~\ref{related-works}) to provide immediate, less costly
and efficient solutions to author disambiguation. In this work, our goal
is to explore and demonstrate how both approaches can coexist and benefit from
each other.  In particular, we study how labeled data obtained through manual
curation (either centralized or crowdsourced) can be exploited (i) to learn an
accurate classifier for identifying corefering authors, (ii) to guide the
clustering of scientific publications by distinct authors in a semi-supervised
way and finally (iii) to match the resulting clusters with existing profiles.
Our analysis of parameters and features on this large training set reveal
that... \glnote{To be completed with actual results.}

@@

Depending on the type and amount of data available, machine learning approaches
for author disambiguation are usually inventorized into three classes of
algorithms \citep{smalheiser2009author,levin2012citation}.

@@
